Size Matters: free-energy calculations of amino acid adsorption over pristine graphene

### 2023-03-10 paper/main.tex
Updated to follow overleaf

\documentclass[twoside,twocolumn,9pt]{article}

\usepackage[super,sort&compress,comma]{natbib}
\usepackage[left=1.5cm, right=1.5cm, top=1.785cm,
bottom=2.0cm]{geometry}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{{./figures}{../figures}}

%\usepackage{epstopdf}
\usepackage{epstopdf}
\author{Mateo Barria-Urenda, José Antonio Gárate}
\title{Thermodynamic characterization of the adsorption of amino acids
  onto pristine graphene and graphene oxide}
\date{}

@@

\abstract{}

@@

% Context

First synthesized in 2004 \cite{Novoselov_2004}, graphene has since
experienced an explosive growth in interest \cite{Randviir_2014}.  For
pristine graphene, it's interactions with other particles are mainly
due to Van der Waals forces and $\pi-\pi$ stacking. \cite{Zuo_2012} As
pristine graphene is chemically inert \cite{Eftekhari_2017} a
classical description of it is expected to suffice in simulations.
Multiple molecular dynamics studies on the interactions between
proteins and carbon based nanoparticles (CBNs) --such as graphene--
have been published \cite{Zheng_2003, Ge_2011, Zuo_2012, Chong_2015,
  Duan_2015, Shityakov_2015, Al_Qattan_2018,Puigpelat_2019,
  Gonz_lez_Durruthy_2020, Li_2020}

% Need

% Task

% Work

@@

\subsection{Umbrella Sampling}

\subsection{Well Tempered Multiple Walk Metadynamics}

\subsection{Molecular Dynamics Simulation}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{/home/mbarria/Dropbox/Papers/graphene_adsorption_paper/figures/Pristine_System.png}}
\caption[]{\label{fig:system-pristine} Example of simulated pristine
  graphene systems. An amino acid (in this case Phenylalanine) above a
  graphene layer (in gray) inside a periodic water box.}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{/home/mbarria/Dropbox/Papers/graphene_adsorption_paper/figures/Oxidized_System_and_Unit.png}}
\caption[]{\label{fig:system-oxidized} a) Example of simulated oxidized
  graphene systems. An amino acid (in this case Phenylalanine) above a
  graphene layer (C in gray, O in red, H in white) inside a periodic
  water box. b) Segment of the oxidized graphene layer depicting the
  oxidation pattern. (C in gray, O in red, H in white)}
\end{figure}

\subsection{Quantum Mechanics / Molecular Mechanics}

\section{Results}

\begin{figure*}[htbp]
\centerline{\includegraphics[width=\textwidth]{/home/mbarria/Dropbox/Papers/graphene_adsorption_paper/figures/Differences_by_names_and_weight.png}}
\caption[]{\label{fig:differences} Differences of adsorption over a
  pristine graphene layer for free energies ($\Delta A_{ads}$),
  energies ($\Delta E_{ads}$) and entropies ($T \Delta S_{ads}$) for
  all proteinogenic amino acids. Top row (a, c, and e):
  $\Delta A_{ads}$, $\Delta E_{ads}$, and $T \Delta S_{ads}$, arranged
  by amino acid classification on the basis of side-chain
  interactions. Amino acids are labeled on the x-axis by their three
  letter code. Bottom row (b, d, and f): $\Delta A_{ads}$,
  $\Delta E_{ads}$, and $T \Delta S_{ads}$, as a function of molecular
  weight of the amino acid. Amino acids are labeled by their markers
  with their one letter code. In all cases amino acids are in a
  neutral state unless marked with a positive ($+$) or negative ($-$)
  sign.}

\end{figure*}

\begin{figure*}[htbp]
\centerline{\includegraphics[width=\textwidth]{/home/mbarria/Dropbox/Papers/graphene_adsorption_paper/figures/Differences_by_names_and_weight_25o.png}}
\caption[]{\label{fig:differences} Differences of adsorption over an
  oxidized graphene layer for free energies ($\Delta A_{ads}$),
  energies ($\Delta E_{ads}$) and entropies ($T \Delta S_{ads}$) for
  select amino acids. All values are relative to Glycine's, which is
  set to 0. Top row (a, c, and e): $\Delta A_{ads}$,
  $\Delta E_{ads}$, and $T \Delta S_{ads}$, arranged by amino acid
  classification on the basis of side-chain interactions. Amino acids
  are labeled on the x-axis by their three letter code. Bottom row (b,
  d, and f): $\Delta A_{ads}$, $\Delta E_{ads}$, and
  $T \Delta S_{ads}$, as a function of molecular weight of the amino
  acid. Amino acids are labeled by their markers with their one letter
  code. In all cases amino acids are in a neutral state unless marked
  with a positive ($+$) or negative ($-$) sign.}

\end{figure*}

\subsection{Adsorption Free Energy}

\subsection{Adsorption Entropy}

\subsection{Adsorption Energy}

\subsection{Diffusion}

>>>>

%\documentclass[twoside,twocolumn,9pt]{article}
\documentclass[journal=jcisd8,manuscript=article,layout=twocolumn]{achemso}

%\usepackage[super,sort&compress,comma]{natbib}
\usepackage{natbib}
%\usepackage[left=1.5cm, right=1.5cm, top=1.785cm,bottom=2.0cm]{geometry}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{{./figures}{../figures}}
\usepackage{xcolor}

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}

\usepackage{epstopdf}

%% color
\usepackage{ulem}

\SectionNumbersOn

\title{Size Matters: free-energy calculations of amino acid adsorption over pristine graphene}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Mateo Barria-Urenda}
\affiliation{Centro Interdisciplinario de Neurociencia de Valparaíso.}
\alsoaffiliation{Doctorado en Ciencias, Mención Biofísica y Biología Computacional, Facultad de Ciencias, Universidad de Valparaíso}
\alsoaffiliation{Millennium Nucleus in NanoBioPhysics (NNBP)}
\author {Alvaro Ruiz-Fernandez}
\affiliation{Centro Científico y Tecnológico de Excelencia Ciencia \& Vida.}
\author{Carlos Gonzalez}
\affiliation{Millennium Nucleus in NanoBioPhysics (NNBP)}
\alsoaffiliation{}
\author{Chris Oostenbrink}
\affiliation{Institute for Molecular Modeling and Simulation, University of Natural Resources and
Life Sciences, 1190 Vienna, Austria.}

\author{Jose Antonio Garate}
\affiliation{Facultad de Ingeniería, Arquitectura y Diseño, Universidad San Sebastián, Bellavista , Santiago, Chile.}
\alsoaffiliation{Millennium Nucleus in NanoBioPhysics (NNBP)}
\alsoaffiliation{Centro Científico y Tecnológico de Excelencia Ciencia \& Vida.}
\alsoaffiliation{Centro Interdisciplinario de Neurociencia de Valparaíso.}

\email{jose.garate@uss.cl}

@@

\abstract{
There is a still growing interest in graphene interactions with proteins; both for its possible biological applications and due to concerns over detrimental effects at the cellular level.
As with any process involving proteins, an understanding from amino acid composition is desirable.
In this work we systematically studied the adsorption process of amino acids onto pristine graphene via rigurous free-energy calculations.
We characterized the free energy, potential energy, and entropy of adsorption of all proteinogenic amino acids.
The energetic components were further separated into pair interaction contributions.
A linear correlation was found between the free-energy and the Solvent Accessible Surface Area  change  during adsorption($\Delta$SASA$^{ads}$).
Free energies over pristine graphene were compared with adsorption onto graphene oxide (GO), finding an almost complete loss of the favorability of amino acid adsorption onto graphene.
Finally, the correlation with the $\Delta$SASA$^{ads}$ was used to successfully predict the free-energy of adsorption of penta-L-alanine in two different structural states.
Due to the relative ease of calculating the ($\Delta$SASA$^{ads}$ compared to free-energy calculations, it could prove a cost-effective predictor of the free-energy of adsorption for proteins onto non-polar surfaces.
}

@@

% Context
%https://www.hindawi.com/journals/bmri/2021/5518999/
% http://graphenetimes.com/2009/12/boehms-1961-isolation-of-graphene/

Pristine graphene is a mono-layer molecule formed by a honeycomb lattice of aromatic carbons. First observed by Boehm in 1964\cite{Boehm_1962} and synthesized in 2004, \cite{Novoselov_2004} graphene has since experienced an explosive growth in interest. \cite{Randviir_2014}
In particular for graphene and its derivatives, many biological applications have been proposed such as drug-delivery, cancer treatment, antiseptics, neuronal stimulation, cellular imaging among others. \cite{Sun2008,Krishnamoorthy2012,Yue2015}
Moreover, several studies have shown that graphene strongly interacts with membranes and extracellular proteins\cite{Zhang2013,Chong2015,Puigpelat_2019} and in some cases produces detrimental effects at the cellular level. \cite{Duan_2015}
The latter has generated concerns over the biological safety of graphene and similar materials. \cite{Zhou2014,Lalwani_2016}

In a biological context, the majority of the interactions of graphene with bio-molecules occur in an aqueous environment, in particular for extracellular proteins.
Therefore, the water-graphene interaction is an essential component to understand the thermodynamics of the protein-graphene interface; many experimental and theoretical studies have explored the water-graphene interaction and, in general, pristine graphene, graphite and its derivatives are considered to be hydrophobic materials, \cite{Shih2012,Rafiee2012,Taherian2013} even though recent studies have presented data that challenges this general consensus, indicating that pristine graphene and graphite are mildly hydrophilic. \citep{Li2013,Ashraf2014,Mucksch2015,Parobek2015,Hong2016,Kozbial2017,Prydatko2018}
Not withstanding, graphene is not water soluble, limiting its usage in the aqueous phase.
In this regard, graphene oxide (GO) a functional derivative of graphene is more suitable for biological applications as --due to the presence of carboxyl, hydroxyl and epoxy groups-- it is able to disperse in water. \cite{Gomez-Navarro2010,Rhazouani2021}

Graphene is chemically inert and electroneutral. \cite{Eftekhari_2017}
Thus, the binding of proteins to graphene is a physisorption process dominated by weak electrostatics.
This is a result of dipolar interactions be they in the form of permanent, induced or fluctuating dipoles.
The latter are the well known London dispersion forces.
Given the apolar nature and the presence of a delocalized $\pi$-electron cloud in graphene, any molecular model of it should, in principle, include dispersion interactions and induced polarizability. \cite{Hong2016}

Classical representations of graphene in the context of (bio)molecular dynamics simulations (MD), in their majority, model it as a fixed array of neutral sp$^2$ carbons with dispersion interactions included via a mean-field approximation employing the traditional 12-6 Lennard-Jones (LJ) formulation. \cite{Zuo_2012,Ho2014,Jimenez2014,Striolo2016,Wloch2017}
Only lately has induced polarizability, specially for the water-graphene interface, been introduced. \cite{Ho2013,Misra2017}
In a recent work, \cite{Escalona_2022} we have shown that the charge-on-spring models\cite{Yu2003} produced small effects on the water dipolar alignment with the graphene surface (at the first water layer) and water contact angle (WCA) estimations which, anyhow, were more dependent on the water model.
Indeed, the use of the SPC water model in conjunction with the GROMOS53A6 parameter set \cite{Oostenbrink_2005} rendered WCA below 90º, reflecting the above mentioned mildly hydrophilic nature of graphene; including induced polarizability (only for graphene) reduced WCA estimates by around 10º.
Overall, calculations for other molecules only using LJ potentials for graphene are in fair agreement with quantum mechanical computations \cite{Yang2012,Kamel2020} and experimental data on adsorption free energies of small organic species over multi-walled carbon nanotubes. \citep{Comer2015}
In fact, a recent comparison between non-polarizable and polarizable force fields (FF) showed that the former outperformed in predicting experimental binding affinities towards graphene. \cite{Poblete2017}
Consequently, a classical description employing bio-molecular FF for the protein-graphene interaction is expected to suffice.

Multiple MD studies on the interactions between proteins and carbon based nanoparticles (CBNs) --such as graphene-- have been published. \cite{Zheng_2003, Ge_2011, Zuo_2012, Chong_2015,Duan_2015, Shityakov_2015, Mucksch2015, Al_Qattan_2018,Puigpelat_2019,Gonz_lez_Durruthy_2020, Li_2020}
Proteins as a whole, regardless of their physico-chemical nature, do adsorb onto pristine graphene and similar materials.
The latter has been confirmed by atomic force and surface plasmon resonance experiments. \cite{Duan_2015,Chong2015}
The adsorption molecular mechanism has been associated with interactions of hydrophobic and aromatic residues, nonetheless these have, in general, ignored the contribution of water removal from the graphene surface. \cite{Liu2015}
%Quite interestingly, carbon-nanotubes are normally dispersed with polyarginines --a positively charged amino acid-- evidencing that there are other factors to be considered than just the hydrophobic or aromatic nature of the residue. (buscar referencia)
In this regard it becomes important to systematically study the binding affinity of all proteinogenic amino acids to evaluate the contribution of each side chain on the adsorption strength towards graphene.
Up to our knowledge, only few articles have carried out these type of studies via MD simulations \cite{Hughes2015,Welch2015,Dasetty2019} and no quantitative experimental data on all amino acids has been reported so far.

In this work we have systematically calculated via rigorous free energy calculations, the free-energy, energy and entropy of adsorption for all 20 proteinogenic amino acids onto pristine graphene in the context of the GROMOS parameter set.\cite{Oostenbrink_2005}
For charged amino acids in water, we also performed calculations for their neutral versions to evaluate the influence of full charges.
The effect of partial graphene oxidation for representative amino acids was also investigated.
Finally, we evaluated how these individual adsorption free energies can serve to predict the adsorption free energy of a penta-L-alanine peptide for its extended and helical configuration.

Overall and as shown by others, \cite{Hughes2015,Welch2015,Dasetty2019} all amino acids present favorable adsorption free energies towards pristine graphene, with the aromatic ones exhibiting the most favorable values.
Quite interestingly, we found a strong correlation between the molecular weight and binding affinity for uncharged amino acids.
This correlation was further confirmed, as there was also a linear dependence between their adsorption strength and the interaction surface ($\Delta$SASA of the unbound and bound states), revealing that the adsorption process towards pristine graphene is preceeded by a desolvation and a dewetting phenomenon.
Accordingly, partial graphene oxidation largely removed the correlation between size and binding strength.
At last, the penta-L-alaline extended configuration presented stronger affinities towards pristine graphene when compared to their helical counterpart, a fact that was predicted by the linear $\Delta$SASA$^{ads}$ models obtained from the single amino acid calculations.
Our results highlight the role of interfacial water on the binding process towards hydrophobic or mild hydrophylic surfaces and suggest that the $\Delta$SASA$^{ads}$ of exposed residues in conjunction with single amino acid affinities may serve as predictors for binding affinities in proteins adsorbed onto hydrophobic surfaces.

@@

\subsection{Free energy, energy and entropy calculations}

The Helmholtz free energy of adsorption of an amino acid onto a graphene layer ($\Delta A^{ads}$) can be obtained from the difference in free energy of the near and free states:

\begin{equation}
\label{eq:Adsorption}
\Delta A^{ads} = A^{near} - A^{free}
\end{equation}
%
where the near and free states are defined based on the projection of $A$ along a reaction coordinate $\xi(\mathbf{r})$.
This projection of $A$ is also known as the Potential of Mean Force (PMF).
For our simulations, $\xi(\mathbf{r})$ was defined as the distance between the graphene layer and the $\alpha-$carbon of the amino acid projected onto a vector normal to the graphene layer.
For a graphene layer prepared along the xy plane, this is equivalent to the distance along the z--axis.
The cut-off ranges for the near and far states were defined based on the PMFs along $\xi(\mathbf{r})$; the near state was defined as the PMF valley ($\xi(\mathbf{r})\leq 0.6 [nm]$) while the free state encompassed the region in which the PMF reaches at a plateau ($1.1 [nm] \leq\xi(\mathbf{r})$).
For more details see Figure S1.
In the case of the single water molecule adsorption, the distance between the oxygen atom and the graphene layer projected along the graphene normal was chosen to define $\xi$.

After obtaining the PMF$(\xi)$, the $\Delta A^{ads}$ was computed by:

\begin{equation}
\label{eq:Aads-from-PMF}
\Delta A^{ads} = - k_B T \ln \left(\frac{\sum_{j \in near} e^{-\beta~\mathrm{PMF}(\xi_j)}}{\sum_{k \in free} e^{-\beta~\mathrm{PMF}(\xi_k)}}\right)
\end{equation}
%
with $\beta=1/k_BT$. Where $k_B$ is Boltzmann's constant, $T$ is the temperature, and the sums run over the {near} and far segments of the computed points of the PMF.

Umbrella sampling \cite{Torrie_1977} via a staging procedure was used to obtain the PMF along the reaction coordinate by sampling multiple windows along $\xi$.
Each sample was obtained from a simulation where the Hamiltonian includes the following harmonic potential:

\begin{equation}
\label{eq:umbrella-potential}
\mathcal{V}^{US}(\mathbf{r}, k^{US}, \xi^0) = \frac{1}{2}k^{US}[\xi^0
- \xi(\mathbf{r})]^2
\end{equation}
%
where $\xi^0$ is the center of the sampling window and $k^{US}$ is an harmonic constant.
For the simulations of amino acids over pristine graphene $\xi^0$ took values from 0.3 to 1.5 nm in 0.05 nm steps, while $k^{US}$ took values from 1000 to 12000 [$kJ \cdot mol^{-1} \cdot nm^{-2}$] adjusted to ensure a good overlap in the sampling of different windows while retaining a high sampling of the window's center.
On average, each simulated window  was run for 25 ns.

The biased samples of $\xi$ obtained from these windows were used to obtain the unbiased probabilities using an implementation of the Weighted Histogram Analysis Method (WHAM) \cite{Kumar_1992, Kumar_1995} developed by Grossfield.\cite{WHAM}
Errors for the PMF profiles were estimated using the software-provided Monte-Carlo bootstrap method, with a count of 1000 trials per PMF.
Grossfield's implementation of WHAM outputs --in addition to the free energy/probability along $\xi$-- the free energy of each window, $F_i$.
In this way, we can unbias the average of any property $X$, as a function of $\xi$:

\begin{equation}
\label{eq:WHAM-unbias}
\langle X(\xi') \rangle = \frac{\sum_i^{S} \sum_j^{N_i} X_{i, j}
  e^{-\beta[F_i - \mathcal{V}^{US}(\mathbf{r_j})]}
  \delta \xi_{i, j}}{\sum_i^{S} \sum_j^{N_i}
  e^{-\beta[F_i - \mathcal{V}^{US}(\mathbf{r_j})]}
  \delta \xi_{i, j}}
\end{equation}
%
where $\xi'$ is the center of a US window, S is the total number of simulations, $N_i$ is the total number of points sampled in the $i$-th simulation, and $\delta \xi_{i, j}$ is the delta function that outputs $1$ if the $j$-th frame of the $i$-th simulation is within the bin corresponding to $\xi'$, and 0 otherwise.
Eq.~\ref{eq:WHAM-unbias} amounts to an average of all points that fall within the $\xi'$ bin, weighted by the penalty of the applied harmonic restraint. Accordingly, the total average potential energies $\langle E(\xi')\rangle$ for each of the US windows were unbiased via Eq.\ref{eq:WHAM-unbias} and then for the near and free states, these energies were weighted by the relative probabilities of $\xi$ being within  that region to obtain the average potential energy of adsorption  ($\Delta E^{ads}_{pot}$):

 \begin{align}
    \label{eq:Eads}
    \Delta E^{ads}_{pot} &= \frac{\sum_{j \in near} \langle E(\xi_j)\rangle p(\xi_j)}{\sum_{j \in near} p(\xi_j)} \nonumber \\
    &- \frac{\sum_{k \in free} \langle E(\xi_k)\rangle p(\xi_k)}{\sum_{k \in free} p(\xi_k)}
\end{align}
%
where p($\xi_i$) is the probability of $\xi(\mathbf{r})$ being within the bin corresponding to $\xi_i$.
This was estimated from the PMF profiles via the following equation:

\begin{equation}
    p(\xi_i) = \frac{e^{-\beta \mathrm{PMF}(\xi_i)}}{\sum_{l} e^{-\beta \mathrm{PMF}(\xi_l)}}
\end{equation}
%
where $l$ sums over the entirety of the PMF. An error estimation of $\Delta E^{ads}_{pot}$ was obtained via bootstrapping.
In detail, re-sampling was done for the entirety of the US simulations and each sample was used to obtain a new PMF, from which a new $\langle E(\xi)\rangle$ and $\Delta E^{ads}_{pot}$ were obtained.
In total, 1000 bootstrap samples were obtained for each $\Delta E^{ads}_{pot}$ calculation. To reduce error estimates, we extended simulations in the minimal value of the PMF valley and in a window further away from the surface for an extra 100 ns each.

The integration of $A(\xi(\mathbf{r}))^{near}$ will be independent of the boundary as it will be dominated by the free-nergy minimum; this also means that the energy and entropy will also be independent of this cut-off distance. On the other hand, $A(\xi(\mathbf{r}))^{free}$ will depend on the choice of the bounds  as we are restricting the molecule to a certain volume ($V_{free}$). This is a purely entropic effect  due to the use of a cut-off for long-range interactions, thus $E_{pot}^{free}$ will be constant and volume independent. Thus the final term to add will be the standard state correction:

\begin{equation}
\label{eq:Standard}
\Delta A^{0} =-k_BT \ln \frac{V_{free}}{V_{0}}
\end{equation}
%
where $V_0$ is the standard state volume (1.61
 $nm^3$).

Finally, the (standard) entropy of adsorption  ($T\Delta S_{ads}^0$) is then:

\begin{equation}
\label{eq:Entropy}
T\Delta S_{ads}^0 = \Delta E^{ads}_{pot}-(\Delta A^{ads} +\Delta A^{0})
\end{equation}

\subsection{Molecular dynamics simulations}

Molecular Dynamics (MD) simulations were performed using the GROMOS11 simulation engine.\cite{Riniker_2011,Schmid_2012}
The SHAKE algorithm \cite{Ryckaert_1977} was employed to constrain all bonds to their reference values with a relative tolerance of $10^{-4}$, allowing for a time-step of 2 fs using the leapfrog algorithm.\cite{Hockney_1977}
For water, the SPC model was used.\cite{Berendsen_1981}
Non-bonded interactions were computed using a triple range cut-off.
Interactions within the short-range cut-off of 0.8 nm were calculated every time-step, from a pair-list that was generated every 10 fs.  At these time points, interactions in the range of 0.8 and 1.4 nm were also computed which were otherwise kept constant between these updates.
A reaction-field contribution was added to electrostatic interactions approximating a homogeneous medium outside the 1.4 nm long-range cut-off, employing the relative
permittivity of SPC water (61).\cite{Tironi_1995}
All interactions were calculated using the GROMOS 54A8 potential energy function, with all graphene atoms modeled as neutral sp$^2$ carbons.\cite{Reif_2012}
This is the same model and parameters used in previous work.\cite{Escalona2022}
After a steepest-descent minimization to remove bad contacts, all velocities were randomly assigned from a Maxwell-Boltzmann distribution at 298 K.
All simulations were run coupled to thermostats using the weak-coupling algorithm.\cite{Berendsen_1984}
The solute, graphene and solvent atoms were independently coupled to different heat baths.
Additionally, the graphene layer was coupled to separate baths for the  regulation of its center of mass motion and its rotational and internal degrees of freedom, respectively.

A periodic graphene layer with a bond-length of $0.139$ nm and edge lengths of  $4.94$  and $4.87$ nm along the X and Y dimensions, respectively, were generated with VMD \cite{Humphrey_1996} using the nanotube builder plugin.
This layer was placed along the X-Y plane in a $5.1 \times 5.0 \times 5.0$ nm$^3$ box; the lengths of the sides of the box along the X and Y axes were chosen to account for periodic bonds between the edges of the graphene layer.
For all simulations the graphene layer was positionally restrained.
The solute was then placed above the graphene layer, and the box filled with pre-equilibrated SPC water maintaining a minimum distance  of $0.23$ nm.
PMF calculations for all  proteinogenic amino acids were performed with their C and N-terminal in their neutral forms.
Calculations for the adsorption of a single water molecule were carried out as well.
For Arg, Lys, His, Glu, and Asp, both the charged and neutral species were simulated.
We also repeated these calculations for certain amino acids (and a water molecule) over a partially oxidized graphene layer.
Partial graphene oxidation was modeled by adding  hydroxyl groups to the 25\% of carbon atoms in regularly-spaced opposite-oriented (trans isomeric) pairs.
The amino acids in question were: Arg(+), Glu(-), Asn, Gly, Ile, and Phe.
For these systems, the boundaries of the adsorbed and free states were instead defined as $\xi(\mathbf{r}) \leq 0.65~[nm]$ and $\xi(\mathbf{r}) \geq 1.5~[nm]$, respectively, and $\xi^0$ took values in range of 0.3 to 2.0 $[nm]$.
A depiction of the pristine and oxidized grapehne system is shown in Fig.\ref{fig:system-aminoacid}.
Lastly, PMF estimations for a penta-L-alanine peptide (with the C and N-terminal in their neutral forms) in the extended and helical configurations were also executed.
For these systems, as shown in Fig.~\ref{fig:system-peptide}, distance restraints between consecutive $\alpha$-carbons were utilized to maintain both configurations and thus avoid structural changes as poly-alanines at room temperature spontaneously transition between folded and unfolded states.\cite{Garate2019}
For the penta-L-alanine peptide, its third  (thus central) $\alpha-$carbon was chosen to define $\xi$.
For both peptide systems, the boundaries of the adsorbed and free states were instead defined as $\xi(\mathbf{r}) \leq 0.9~[nm]$ and $\xi(\mathbf{r}) \geq 1.4~[nm]$, respectively, and $\xi^0$ took values in a range of 0.3 to 2.0 $[nm]$.

\begin{figure}[htbp]
\centerline{\includegraphics[width=.9\columnwidth]{Fig1.pdf}}
\caption[]{\label{fig:system-aminoacid} Example of simulated systems for single amino acids. (a) A phenylalanine  placed above a layer of pristine
  graphene inside a water-filled orthorombic box. (b) A phenylalanine molecule placed above a layer of partially oxidized graphene. Carbon atoms are depicted in gray whithin  the graphene layer, and cyan in the solute. Oxygen atoms and hydrogen atoms are depicted in red and white, respectively. $\xi$ is defined as the distance between the C$\alpha$ and the graphene layer projected along the $z$ axis.}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=.9\columnwidth]{Fig2.pdf}}
\caption[]{\label{fig:system-peptide} Simulated penta-L-alanine systems over pristine graphene. (a) Penta-L-alanine  in its $\alpha$-helical form; (b) Penta-L-alanine in its extended configuration. In both cases, harmonic distance restraints were used to keep the peptide in its respective configuration. Carbon atoms are depicted in gray. Penta-L-alanine molecules are depicted in yellow. $\xi$ is defined as the distance between the third C$\alpha$ and the graphene layer projected along the $z$ axis}
\end{figure}

\section{Results and discussion}

In this section we present the computed $\Delta A^{ads}$ for all 20 proteinogenic amino acids over pristine graphene.
We also investigate the energetic and entropic components of $\Delta A^{ads}$, paying attention to the different energy terms of the  $\Delta A^{ads}$.
We explore how these terms correlate with amino acid size and solvent exposure.
Additionally, we evaluate how partial oxidation of the graphene surface affects $\Delta A^{ads}$ for representative amino acids.
Finally, the possibility of employing the single amino-acids $\Delta A^{ads}$ to predict affinities of a penta-L-alanine peptide in its unfolded and $\alpha$-helical configurations towards graphene is evaluated.

\subsection{Adsorption free energy of amino acids over pristine graphene is correlated with size}

\begin{table}[htbp!]
\centering
\caption{\label{table:freeEnergy} Adsorption free energies ($\Delta A^{ads}0$), energies ($\Delta E^{ads}_{pot}$) and standard entropies ($T \Delta S^{ads}$)  in $[kJ \cdot mol^{-1}]$ for all amino acids (and a water molecule) over pristine graphene. The Side-chain charged state is denoted after the amino acid's name when relevant.}
\resizebox{\linewidth}{!}{%
\begin{tabular}{lrrr}
\toprule
%Name & $\Delta A^{ads}$ & $\Delta E^{ads}$ & $T \Delta S^{ads}$ \\ \hline
\multirow{2}{*}{Name}  & \multicolumn{1}{r}{$\Delta A^{ads}$} & \multicolumn{1}{r}{$\Delta E^{ads}_{pot}$ } & \multicolumn{1}{r}{$T \Delta S^{ads}$}  \\
\cline{2-4}            & \multicolumn{3}{c}{$[kJ \cdot mol^{-1}]$} \\ \hline
Alanine           & $-9.6  \pm 0.1$ & $-13.0 \pm 1.6$  & $-3.4  \pm 1.6$  \\
Arginine          & $-20.7 \pm 0.0$ & $-42.4 \pm 3.4$  & $-21.7 \pm 3.4$  \\
Arginine (+)      & $-18.5 \pm 0.0$ & $-53.3 \pm 5.8$  & $-34.8 \pm 5.8$  \\
Asparagine        & $-13.9 \pm 0.0$ & $-24.8 \pm 1.8$  & $-10.9 \pm 1.8$  \\
Aspartate (-) & $-9.5  \pm 0.0$ & $-11.0 \pm 1.6$  & $-1.5  \pm 1.6$  \\
Aspartic Acid     & $-15.9 \pm 0.0$ & $-25.9 \pm 2.1$  & $-10.0 \pm 2.1$  \\
Cysteine          & $-13.3 \pm 0.0$ & $-16.5 \pm 1.7$  & $-3.2  \pm 1.7$  \\
Glutamate (-) & $-7.5  \pm 0.0$ & $-6.3  \pm 1.5$  & $1.2   \pm 1.5$  \\
Glutamic Acid     & $-17.2 \pm 0.0$ & $-28.3 \pm 3.1$  & $-11.1 \pm 3.1$  \\
Glutamine         & $-16.4 \pm 0.0$ & $-24.1 \pm 1.5$  & $-7.6  \pm 1.5$  \\
Glycine           & $-8.0  \pm 0.0$ & $-16.2 \pm 1.4$  & $-8.3  \pm 1.4$  \\
Histidine (+)     & $-12.3 \pm 0.1$ & $-39.5 \pm 1.6$  & $-27.2 \pm 1.6$  \\
Histidine         & $-18.2 \pm 0.0$ & $-25.5 \pm 1.4$  & $-7.2  \pm 1.4$  \\
Isoleucine        & $-14.3 \pm 0.1$ & $-16.4 \pm 1.5$  & $-2.1  \pm 1.5$  \\
Leucine           & $-15.3 \pm 0.0$ & $-18.3 \pm 1.9$  & $-3.1  \pm 1.9$  \\
Lysine            & $-17.7 \pm 0.0$ & $-24.9 \pm 2.6$  & $-7.1  \pm 2.6$  \\
Lysine (+)        & $-11.4 \pm 0.1$ & $-27.4 \pm 1.6$  & $-16.0 \pm 1.6$  \\
Methionine        & $-19.4 \pm 0.0$ & $-38.3 \pm 6.2$  & $-18.9 \pm 6.2$  \\
Phenylalanine     & $-18.4 \pm 0.1$ & $-27.2 \pm 1.6$  & $-8.8  \pm 1.6$  \\
Proline           & $-6.0  \pm 0.0$ & $-24.1 \pm 1.5$  & $-18.1 \pm 1.5$  \\
Serine            & $-11.2 \pm 0.0$ & $-16.4 \pm 1.6$  & $-5.2  \pm 1.6$  \\
Threonine         & $-13.0 \pm 0.0$ & $-16.7 \pm 1.6$  & $-3.6  \pm 1.6$  \\
Tryptophan        & $-26.2 \pm 0.0$ & $-33.2 \pm 1.4$  & $-7.1  \pm 1.4$  \\
Tyrosine          & $-27.5 \pm 0.0$ & $-30.7 \pm 1.6$  & $-3.2  \pm 1.6$  \\
Valine            & $-12.6 \pm 0.0$ & $-11.5 \pm 1.6$  & $1.1   \pm 1.6$  \\
Water             & $3.3   \pm 0.1$ & $-3.6  \pm 2.8$  & $-6.9  \pm 2.8$  \\
\bottomrule
\end{tabular}
}
\end{table}

In the second column of Table \ref{table:freeEnergy} we present the $\Delta A_{ads}^0$ from the aqueous phase for all 20 proteinogenic amino-acids (and water) over pristine graphene computed via eq.\ref{eq:Adsorption}; the PMF profiles obtained via umbrella sampling are shown in Fig.S1. For Arg, Lys, His, Glu, and Asp, their side-chains in the neutral and charged states were computed as well.

All amino acids, not withstanding their chemical nature, present favorable $\Delta A^{ads}$ towards pristine graphene. This is in line with previous works by \citet{Dasetty2019} in which they tested several other force-fields, but their values are consistently deeper when compared to the ones reported in Table \ref{table:freeEnergy}.
Comparing free-energies require standard state corrections, as the volume for the unbound states may differ; thus  we corrected all $\Delta A^{ads}$ values  via eq. \eqref{eq:Standard}, which are shown in Figs. S10a and S10b. Notwithstanding, the differences with the values reported by \citet{Dasetty2019} still remain.

In this previous study all amino-acids were capped with acetyl and n-methyl amide groups \cite{Dasetty2019}, while in our case we just employed the uncharged N-terminal and C-terminal (NH$_2$ and COOH, respectively).
Consequently and to isolate the effects of the side-chains, in Fig. \ref{fig:AadsCorre}A we present the $\Delta A^{ads}$ reduced by the $\Delta A^{ads}$ of Gly.
When these reduced values are then compared our values exhibit a similar trend) and converge with previously repor ted data.\cite{Dasetty2019}
This comparison is displayed in Fig.S10 of the supplementary material, where the data shown in Fig.3b of \citet{Dasetty2019} is presented along with that of Fig.\ref{fig:AadsCorre}A of this study.

\begin{figure}[htbp]
\centerline{\includegraphics[width=.95\columnwidth]{Fig3.pdf}}
\caption[]{\label{fig:AadsCorre} Reduced adsorption free energies (i.e. with $\Delta A^{ads}$ of Gly set to $0$) for all amino acids. Unless otherwise noted, amino acids are in their neutral state. Charged species are indicated with a $+$ or $-$. (a) Reduced free energies ordered by the side-chain nature;  apolar (blue), aromatic (cyan), negative (green), polar (yellow), and positive (red). (b) $\Delta A^{ads}$ as a function of the molecular weight. (c) $\Delta A^{ads}$ as a function of $\Delta \textrm{SASA}^{ads}$. In (b) and (c) a linear regression is shown with a dotted line, and its respective $r^2$ value. This linear regression does not include amino acids in their charged state.}
\end{figure}
